Eurovision Song Contest: semifinal 1, 2009

### 2020-02-20 inst/eurovision.Rmd
slight rewording and reworking

Variable `points` gives the number of points awarded, under Eurovision
rules, to voters' first, second, third, etc choice.  Points for any
competitor are added and the winner is the competitor with the most
points.  However, in this model, the numerical values themselves do
not affect the likelihood function; only the order of the voters'
preferences matters.

@@

Now, take the first row of 'preference'.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.

We need to convert matrix `preference` into a likelihood function,
here `euro2009`:

>>>>

Variable `points` gives the number of points awarded, under Eurovision
rules, to voters' first, second, third, etc choice.  Points for any
competitor are added and the winner is the competitor with the most
points.  However, in the `hyper2` model, the numerical values
themselves do not affect the likelihood function; only the order of
the voters' preferences matters.  The following R idiom translates
`wiki_matrix` into a form suitable for analysis with `hyper2`:

@@

Now, take the first row of `preference`.  This represents the votes
cast *by* (sic) Montenegro (`"ME"`).  Their favourite was [last
column] Bosnia & Herzegovina (`"BA"`), who they gave rank 1 to.  Their
second favourite was Macedonia, their third was Turkey, and so on.
They were not allowed to vote for themselves, which is why the first
column is `NA`.  So the order was: `BA` (1), `MK` (2) , `TR` (3), `IS`
(4), `RO` (5), `IL` (6), `AM` (7), `FI` (8), `BY` (9), `MT` (10).  The
other countries (`CZ`, `BE`, `SW`, `CH`, `BG`, `PT`) did not receive
any points from Montenegro.  We need to convert matrix `preference`
into a likelihood function, here `euro2009`:
